---
title: "Homework 7"
author: "Casey Gibson"
date: "4/2/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem 1
$$y_t = \rho y_{t-1} + \epsilon_t$$
$$\gamma_k  = Cov(y_{t-k},y_t) = \rho^k\gamma_0$$
$$y_{t-k}\cdot y_t = y_{t-k} \cdot \rho y_{t-1} +y_{t-k}\cdot \epsilon_t$$
$$E(y_{t-k}\cdot y_t) =\rho E(y_{t-k} \cdot y_{t-1}) +E(y_{t-k})\cdot E(\epsilon_t)$$
by independence.

$$E(y_{t-k} \cdot y_t) = \rho E( y_{t-k} \cdot y_{t-1}) $$
and we can keep going , by writing the right hand side as 

$$E(y_{t-k}\cdot y_{t-1}) = \rho E(y_{t-k} \cdot y_{t-2}) + 0$$
Pluggin this into the above we get 

$$E(y_{t-k}\cdot y_{t}) = \rho^2E(y_{t-k} \cdot y_{t-2})$$
So continuing we see that 
$$E(y_{t-k} \cdot y_t) = \rho^k E( y_{t-k} \cdot y_{t-k}) $$
We know that $$E(y_{t-k}^2) = Var(y_{t-k}) - E(y_{t-k})^2$$
$$E(y_{t-k}^2) = Var(y_{t-k}) - 0^2$$
because $E(y_{t-k}) = 0$.

We also know that 

$$Var(y_{t-k}) = Cov(y_{t-k},y_{t-k}) = Cov(y_t,y_t) =  \gamma_0$$
So plugging back in we see,

$$E(y_{t-k} \cdot y_t) = \rho^k \gamma_0 $$

### Problem 2

```{r}
library(R2jags)
library(rjags)

GetAR <- function(nyears, # length of series
                  rho, sigma, # AR parameters
                  eps0.t = NULL, # innovations (optional)
                  ystart = NULL # starting value y1 (optional)
                  ){
  if (is.null(eps0.t)){
    set.seed(123)
    eps0.t <- rnorm(nyears, 0, 1)
  }
  y.t <- rep(NA, nyears)
  if (is.null(ystart)){
    y.t[1] <- sigma/sqrt(1-rho^2)*eps0.t[1]
  } else {
    y.t[1] <- ystart
  }
  for (t in 2:nyears){
    y.t[t] <- rho*y.t[t-1] + sigma*eps0.t[t]
  }
  return(y.t)
}
rho <- 0.5
sigma <- 1
sigma.y <- 0.5
nyears <- 100
mu.t <- GetAR(nyears, rho, sigma)
set.seed(124)
y.t <- mu.t + rnorm(nyears, 0, sigma.y)


model <- "
model{
mu.t[1] ~ dnorm(0, tau.stat)

tau.stat <- (1-pow(rho,2))/pow(sigma,2)
for (t in 2:(nyears+P)){
 mu.t[t] ~ dnorm(muhat.t[t], tau)
 muhat.t[t] <- rho*mu.t[t-1]
 yhat.t[t] ~ dnorm(muhat.t[t], tau.y)

}
for (t in 1:nyears){
  y.t[t] ~ dnorm(mu.t[t], tau.y)
}


tau <- pow(sigma,-2)
sigma ~ dunif(0,3)
tau.y <- pow(sigma.y,-2)
sigma.y ~ dunif(0,3)
rho ~ dunif(-1,1)
}
"

P <- 20
set.seed(1234)
t.i <- sort(sample(seq(1,100), size = 70))
y.t[-t.i] <- NA
jags.data <- list(y.t = y.t, nyears=nyears,P=P)
parnames <- c("sigma", "rho","mu.t","sigma.y","muhat.t","yhat.t")

mod <- jags(data = jags.data, 
           parameters.to.save= c(parnames),  n.chains = 4, n.burnin = 1000, n.iter = 1000+30000, n.thin =30
,
           model.file = textConnection(model))

```

```{r}
mcmc.array <- mod$BUGSoutput$sims.array
mcmc.list <- mod$BUGSoutput$sims.list
mean(mcmc.array[,,"mu.t[6]"])
quantile(mcmc.array[,,"mu.t[6]"],c(.025,.975))
```


```{r,echo=FALSE,message=FALSE}
AddCIs <- function(CI.low.t, # lower bound for seq.years.t
                   CI.up.t, # upper bound for seq.years.t
                   seq.years.t, col = 1){
  # add CIs to a plot.
  col = adjustcolor(col, alpha.f = 0.1)
  for (t in 2:length(seq.years.t))
    polygon(c(seq.years.t[t-1], seq.years.t[t-1], seq.years.t[t], seq.years.t[t],seq.years.t[t-1]),
            c(CI.low.t[t-1], CI.up.t[t-1], CI.up.t[t], CI.low.t[t], CI.low.t[t-1]),
            col=col, border = NA)
}
mu.st <- mod$BUGSoutput$sims.list[["mu.t"]]
mu.qt <- apply(mu.st, 2, quantile, c(0.025, 0.5, 0.975))
par(lwd = 3, cex.axis = 2, cex.lab = 2, cex.main = 2,mar = c(5,5,1,1), mfrow = c(1,1))
plot(y.t, type = "n", ylab = expression(y[t]), xlab = expression(t))
AddCIs(CI.low.t = mu.qt[1,],CI.up.t = mu.qt[3,], seq.years.t = seq(1, nyears), col = 2)
lines(mu.qt[2,], type = "l", col = 2)
# true AR(1)
lines(mu.t, col = "purple", lty = 2)
points(y.t)
mcmc.array <- mod$BUGSoutput$sims.array

```

We can forecast $\mu_t$ using 

$$\mu_{101}^{(s)} | \mu_{100}^{(s)} , \rho^{(s)},(\delta^2)^{(s)} \sim N(\mu_{100}^{(s)},(\delta^2)^{(s)})$$
and forecast $y_t$ using 
$$y_{101}^{(s)} | \mu_{101}^{(s)}, (\sigma^2)^{(s)} \sim N(\mu_{101}^{(s)},(\sigma^2)^{(s)})$$

and so on as we extend in time. 

```{r,echo=FALSE,message=FALSE}

mu_90_120_mean <- c()
mu_90_120_upper_ci <- c()
mu_90_120_lower_ci <- c()
for (i in 90:120){
  mu_90_120_mean <- c(mu_90_120_mean,mean(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")]))
  mu_90_120_upper_ci <- c(mu_90_120_upper_ci,quantile(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")],.975))
  mu_90_120_lower_ci <- c(mu_90_120_lower_ci,quantile(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")],.025))
}


y_90_120_mean <- c()
y_90_120_upper_ci <- c()
y_90_120_lower_ci <- c()
for (i in 90:120){
  y_90_120_mean <- c(y_90_120_mean,mean(mcmc.array[,,paste(paste("yhat.t[",i,sep=""),"]",sep="")]))
  y_90_120_upper_ci <- c(y_90_120_upper_ci,quantile(mcmc.array[,,paste(paste("yhat.t[",i,sep=""),"]",sep="")],.975))
  y_90_120_lower_ci <- c(y_90_120_lower_ci,quantile(mcmc.array[,,paste(paste("yhat.t[",i,sep=""),"]",sep="")],.025))
}

plot(90:100,y.t[90:100],ylim=c(-3,3),xlim=c(90,120),xlab="Time",ylab="muhat",col='red',sub="Blue=UI,Red=CI")
lines(90:120,mu_90_120_mean,ylim=c(-3,3))
lines(90:120,mu_90_120_upper_ci,lty=2,col='red')
lines(90:120,mu_90_120_lower_ci,lty=2,col='red')
lines(90:120,y_90_120_upper_ci,lty=2,col='blue')
lines(90:120,y_90_120_lower_ci,lty=2,col='blue')

```

```{r}
print ("muhat[101]") 
print (mean(mcmc.array[,,"muhat.t[101]"]))
print (quantile(mcmc.array[,,"muhat.t[101]"],c(.025,.975)))


```

#### c

```{r,echo=FALSE}

### BEFORE 97
mu_90_120_mean <- c()
mu_90_120_upper_ci <- c()
mu_90_120_lower_ci <- c()
for (i in 90:97){
  mu_90_120_mean <- c(mu_90_120_mean,mean(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")]))
  mu_90_120_upper_ci <- c(mu_90_120_upper_ci,quantile(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")],.975))
  mu_90_120_lower_ci <- c(mu_90_120_lower_ci,quantile(mcmc.array[,,paste(paste("muhat.t[",i,sep=""),"]",sep="")],.025))
}


#### AFTER 97
mu97 <- mcmc.array[,,"muhat.t[97]"]
rho.s <- mcmc.list[["rho"]]
sigma.s <- mcmc.list[["sigma"]]
S <- length(rho.s)
P <- 23
mu.sp <- matrix(NA, S, P)
set.seed(1)
mu.sp[,1] <- rnorm(S, rho.s*mu97[97], sigma.s)
for (p in 2:P){
   mu.sp[,p]<-rnorm(S, rho.s*mu.sp[,p-1], sigma.s)
}
mu.qp <- apply(mu.sp, 2, quantile, c(0.025, 0.5, 0.975))
upper_quantile <- mu.qp[3,]
lower_quantile <- mu.qp[1,]
plot(90:100,y.t[90:100],ylim =c(-3,3),xlim = c(90,120),ylab="mu_t",xlab="time")
lines(90:97,mu_90_120_mean)
lines(90:97,mu_90_120_lower_ci,lty=2,col='red')
lines(90:97,mu_90_120_upper_ci,lty=2,col='red')
lines(98:120,colMeans(mu.sp))
lines(98:120,lower_quantile,lty=2,col='red')
lines(98:120,upper_quantile,lty=2,col='red')
abline(v = 98)
text(98, -3, "cutoff", col = "red") 
```

```{r}
print ("muhat[98]")

print (mean(mu.sp[,1]))
print (quantile(mu.sp[,1],c(.025,.975)))

```

The PI's that result from part $b$ are smaller than those of part $c$ because we are forecasting from later on in the time series.
