---
title: "Homework 6"
author: "Casey Gibson"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

### Problem 1 

```{r, echo=FALSE,message=FALSE, results='hide',  fig.height=3}
#-------------------------------------------------------------
# 4_logistic
# Leontine Alkema
# based on data and code examples from Bayesian Biostatistics (Lesaffre and Lawson)
#-------------------------------------------------------------

# settings
#setwd("C:/Users/lalkema/Documents/Bayes Umass/")
figdir = "fig/"
library(rjags)
library(R2jags)

source("/Users/gcgibson/Downloads/functions.R")

#----
# data
#-----
# Description of the study (taken from LL)
# This is a randomized controlled clinical trial comparing two oral treatments for  toenail toenail dermatophyte onychomycosis with either of two oral medications: Itraconazol 250 mg daily (treat=0) or Lamisil 250 mg daily (treat=1). The patients received treatment for twelve weeks and were evaluated at 0, 1, 2, 3, 6, 9 and 12 months. As response we have taken the the binarized degree of onycholysis of a subgroup of 294 patients.
# Reference: De Backer, M. and De Keyser, P. and De Vroey, C. and Lesaffre, E. A 12-week treatment for dermatophyte toe onycholysis: terbinafine 250 mg/day vs itraconazole 200 mg/day - a double-blind comparative trial, British Journal of Dermatology, 134, 16-17, 1996
#dat <- read.table(file = "data/toenailb.txt", header = T)
wells <- read.table ("/Users/gcgibson/Downloads/wells.dat")
dim(wells)
# these are the transformed versions of the variables that will be used 
y.i <- wells$switch
n <- length(y.i)
dist100 <- wells$dist/100
# I just center the distance and arsenic variables from the start
c.dist100 <- dist100 - mean (dist100)
c.arsenic <- wells$arsenic - mean (wells$arsenic)
log.arsenic <- log (wells$arsenic)
c.log.arsenic <- log.arsenic - mean (log.arsenic)
# # not yet used
# educ4 <- educ/4
# c.educ4 <- educ4 - mean(educ4)



#-----
# fitting
model <- "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0
  + b1*c.dist100[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
} # end model

"
# with random slopes
jags.data2 <- list(y.i = y.i, 
                   c.dist100 = c.dist100,  n=n)
parnames2 <- c( "b1", "b0")
# note: error messages may be less informative with jags parallel so go back to using "jags" (the non-parallel version) 
# if you don't understand an error message
mod2 <- jags(data = jags.data2, 
                    parameters.to.save=parnames2, 
                    n.chains = 3, n.burnin = 1500, n.iter = 1500, n.thin = 10, model.file = textConnection(model))
max(mod2$BUGSoutput$summary[, c("Rhat")])
min(mod2$BUGSoutput$summary[, c("n.eff")]) 
which.max(mod2$BUGSoutput$summary[, c("Rhat")])
which.min(mod2$BUGSoutput$summary[, c("n.eff")])
# just some trace plots

PlotTrace("b0", mod2$BUGSoutput$sims.array)
PlotTrace("b1", mod2$BUGSoutput$sims.array)

# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod2$BUGSoutput$sims.array
par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,3))
for (k in 0:1){
 hist(c(mcmc.array[,,paste0("b", k, "")]), freq = F, main = "", xlab = paste0("Coeff ", k))
 curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
}
```

```{r}
print 
print (c("B0",(signif(mean(mod2$BUGSoutput$sims.array[,,"b0"]),3))),signif(quantile(mod2$BUGSoutput$sims.array[,,"b0"],c(.025,.975)),3))


print (c(("B1"),signif(mean(mod2$BUGSoutput$sims.array[,,"b1"]),3),signif(quantile(mod2$BUGSoutput$sims.array[,,"b1"],c(.025,.975)),3)))

```

```{r,results="hide"}
 "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0
  + b1*c.dist100[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
} # end model
"
```


In terms of odds ratios we see that there is a $exp(.305) = 1.36$ expected multiplicative effect on the odds of switching wells for households that are $\bar{d}$ meters away from the nearest safe well. This is because when $d_i = \bar{d}$ , $beta_1$ drops out. We expect to see an effect between $(exp(.231),exp(.378)) = (1.26,1.46)$ $95\%$ of the time. 


Similarly, we expect a $exp(-0.620) = .5379$ multiplicative effect on the odds ratio of switching for every unit increase in distance from the nearest safe well. We expect to see an effect somewhere between $(exp(-0.814),exp(-0.438 )) = (.443,.645)$ $95\%$ of the time. 


### Problem 2

#### a) 
```{r, message=FALSE,echo=FALSE,warning=FALSE,results='hide'}
#-----
# fitting
model <- "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  a0 + a1*c.dist100[i] + a2*c.arsenic[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

a0  ~ dnorm(0.0,1.0E-4)
a1 ~ dnorm(0.0,1.0E-4)
a2 ~ dnorm(0.0,1.0E-4)

} # end model

"
# with random slopes
jags.data2 <- list(y.i = y.i, 
                   c.dist100 = c.dist100, c.arsenic=c.arsenic, n=n)
parnames2 <- c( "a1", "a0","a2")
# note: error messages may be less informative with jags parallel so go back to using "jags" (the non-parallel version) 
# if you don't understand an error message
mod2 <- jags(data = jags.data2, 
                    parameters.to.save=parnames2, 
                    n.chains = 3, n.burnin = 1500, n.iter = 1500+10000, n.thin = 10, model.file = textConnection(model))
max(mod2$BUGSoutput$summary[, c("Rhat")])
min(mod2$BUGSoutput$summary[, c("n.eff")]) 
which.max(mod2$BUGSoutput$summary[, c("Rhat")])
which.min(mod2$BUGSoutput$summary[, c("n.eff")])
# just some trace plots
# PlotTrace("a0", mod2$BUGSoutput$sims.array)
# PlotTrace("a1", mod2$BUGSoutput$sims.array)
# PlotTrace("a2", mod2$BUGSoutput$sims.array)


# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod2$BUGSoutput$sims.array
# par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,3))
# for (k in 0:2){
#  hist(c(mcmc.array[,,paste0("a", k, "")]), freq = F, main = "", xlab = paste0("Coeff ", k))
#  curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
# }

a0_mean <- mean(mcmc.array[,,"a0"])
a1_mean <- mean(mcmc.array[,,"a1"])
a2_mean <- mean(mcmc.array[,,"a2"])

```

```{r,results='hide'}
"
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  a0 + a1*c.dist100[i] + a2*c.arsenic[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

a0  ~ dnorm(0.0,1.0E-4)
a1 ~ dnorm(0.0,1.0E-4)
a2 ~ dnorm(0.0,1.0E-4)

} # end model

"
```

```{r,results='hide'}
"
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0 + b1*c.dist100[i] + b2*c.arsenic[i] + b3*c.arsenic[i]*c.dist100[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)
b3 ~ dnorm(0.0,1.0E-4)

} # end model
"
```

#### b) 


```{r, message=FALSE,echo=FALSE,warning=FALSE,results='hide'}
#-----
# fitting
model <- "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0 + b1*c.dist100[i] + b2*c.arsenic[i] + b3*c.arsenic[i]*c.dist100[i]
  )
  y.i[i] ~ dbern(p.i[i])
}

b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)
b3 ~ dnorm(0.0,1.0E-4)

} # end model

"
# with random slopes
jags.data2 <- list(y.i = y.i, 
                   c.dist100 = c.dist100, c.arsenic=c.arsenic, n=n)
parnames2 <- c( "b1", "b0","b2","b3")
# note: error messages may be less informative with jags parallel so go back to using "jags" (the non-parallel version) 
# if you don't understand an error message
mod2 <- jags(data = jags.data2, 
                    parameters.to.save=parnames2, 
                    n.chains = 3, n.burnin = 1500, n.iter = 1500+10000, n.thin = 10, model.file = textConnection(model))
max(mod2$BUGSoutput$summary[, c("Rhat")])
min(mod2$BUGSoutput$summary[, c("n.eff")]) 
which.max(mod2$BUGSoutput$summary[, c("Rhat")])
which.min(mod2$BUGSoutput$summary[, c("n.eff")])
# just some trace plots
# PlotTrace("b0", mod2$BUGSoutput$sims.array)
# PlotTrace("b1", mod2$BUGSoutput$sims.array)
# PlotTrace("b2", mod2$BUGSoutput$sims.array)
# PlotTrace("b3", mod2$BUGSoutput$sims.array)


# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod2$BUGSoutput$sims.array
# par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,4))
# for (k in 0:3){
#  hist(c(mcmc.array[,,paste0("b", k, "")]), freq = F, main = "", xlab = paste0("Coeff ", k))
#  curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
# }

b0_mean <- mean(mcmc.array[,,"b0"])
b1_mean <- mean(mcmc.array[,,"b1"])
b2_mean <- mean(mcmc.array[,,"b2"])
b3_mean <- mean(mcmc.array[,,"b3"])
```

```{r,message=FALSE}
invlogit <- function(x) {exp(x)/(1 + exp(x))}

plot(x = c.arsenic, y =y.i ,ylab="P[switch]", xlab = "Arsenic Level", main = "Probability of Switching versus Arsenic Level")
curve (invlogit(a0_mean + a1_mean*(1-mean(dist100)) + a2_mean*x), add=TRUE)
curve (invlogit(b0_mean + b1_mean*(1-mean(dist100)) + b2_mean*x + b3_mean*(1-mean(dist100))*x), add=TRUE, col = "red", lty = 2)
legend("bottomright", legend = c("Model 2", "Model 3"), col = c(1,2), lty = c(1,2), cex = .5)


```


We can see that the interaction term accounts for the fact that the importance of arsenic levels decrease as distance increases. That is people who are further away from safe wells consider arsenic less important when making the decision to switch. 

### Problem 4

#### a)
```{r, message=FALSE,echo=FALSE,warning=FALSE,results='hide'}

set.seed(12345)
  # assign households to villages
J <- 300
getj1.i <- c(seq(1,J), sample(size = n-J, x = seq(1,J), replace = TRUE))
  
getj2.i <- sort(getj1.i) # now the households are assumed
  #to be sorted by village

#-----
# fitting
model <- "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0 + b1*c.dist100[i] + b2*c.log.arsenic[i] + b3*c.log.arsenic[i]*c.dist100[i] + alpha.j[getj1.i[i]]
  )
  y.i[i] ~ dbern(p.i[i])
}
for (j in 1:J){
 alpha.j[j] ~ dnorm(0, 1/sigma_alpha^2)
}
b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)
b3 ~ dnorm(0.0,1.0E-4)
sigma_alpha ~ dunif(0,200)

} # end model

"
# with random slopes
jags.data2 <- list(y.i = y.i, 
                   c.dist100 = c.dist100, c.log.arsenic=c.log.arsenic, n=n,getj1.i  = getj1.i, J = J)
parnames2 <- c( "b1", "b0","b2", "b3","alpha.j","sigma_alpha")
# note: error messages may be less informative with jags parallel so go back to using "jags" (the non-parallel version) 
# if you don't understand an error message
mod2 <- jags(data = jags.data2, 
                    parameters.to.save=parnames2, 
                    n.chains = 3, n.burnin = 1500, n.iter = 1500+10000, n.thin = 10, model.file = textConnection(model))
max(mod2$BUGSoutput$summary[, c("Rhat")])
min(mod2$BUGSoutput$summary[, c("n.eff")]) 
which.max(mod2$BUGSoutput$summary[, c("Rhat")])
which.min(mod2$BUGSoutput$summary[, c("n.eff")])
# just some trace plots
#PlotTrace("b0", mod2$BUGSoutput$sims.array)
#PlotTrace("b1", mod2$BUGSoutput$sims.array)
#PlotTrace("b2", mod2$BUGSoutput$sims.array)
#PlotTrace("b3", mod2$BUGSoutput$sims.array)

# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod2$BUGSoutput$sims.array
# par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,4))
# for (k in 0:3){
#  hist(c(mcmc.array[,,paste0("b", k, "")]), freq = F, main = "", xlab = paste0("Coeff ", k))
#  curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
# }

b0_mean <- mean(mcmc.array[,,"b0"])
b1_mean <- mean(mcmc.array[,,"b1"])
b2_mean <- mean(mcmc.array[,,"b2"])
b3_mean <- mean(mcmc.array[,,"b3"])
sigma_alpha_mean_1 <- mean(mcmc.array[,,"sigma_alpha"])

```

```{r, message=FALSE,echo=FALSE,warning=FALSE,results='hide'}

set.seed(12345)
  # assign households to villages
J <- 300
getj1.i <- c(seq(1,J), sample(size = n-J, x = seq(1,J), replace = TRUE))
  
getj2.i <- sort(getj1.i) # now the households are assumed
  #to be sorted by village

#-----
# fitting
model <- "
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0 + b1*c.dist100[i] + b2*c.log.arsenic[i] + alpha.j[getj1.i[i]]
  )
  y.i[i] ~ dbern(p.i[i])
}
for (j in 1:J){
 alpha.j[j] ~ dnorm(0, 1/sigma_alpha^2)
}
b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)
sigma_alpha ~ dunif(0,200)

} # end model

"
# with random slopes
jags.data2 <- list(y.i = y.i, 
                   c.dist100 = c.dist100, c.log.arsenic=c.log.arsenic, n=n,getj1.i  = getj2.i, J = J)
parnames2 <- c( "b1", "b0","b2", "b3","alpha.j","sigma_alpha")
# note: error messages may be less informative with jags parallel so go back to using "jags" (the non-parallel version) 
# if you don't understand an error message
mod2 <- jags(data = jags.data2, 
                    parameters.to.save=parnames2, 
                    n.chains = 3, n.burnin = 1500, n.iter = 1500+10000, n.thin = 10, model.file = textConnection(model))
max(mod2$BUGSoutput$summary[, c("Rhat")])
min(mod2$BUGSoutput$summary[, c("n.eff")]) 
which.max(mod2$BUGSoutput$summary[, c("Rhat")])
which.min(mod2$BUGSoutput$summary[, c("n.eff")])
# just some trace plots
#PlotTrace("b0", mod2$BUGSoutput$sims.array)
#PlotTrace("b1", mod2$BUGSoutput$sims.array)
#PlotTrace("b2", mod2$BUGSoutput$sims.array)

# some priors and posteriors to check that priors weren't informative
mcmc.array <- mod2$BUGSoutput$sims.array
# par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,4))
# for (k in 0:2){
#  hist(c(mcmc.array[,,paste0("b", k, "")]), freq = F, main = "", xlab = paste0("Coeff ", k))
#  curve(dnorm(x,0,sqrt(1/10^(-4))), add = T, col = 2)
# }

b0_mean <- mean(mcmc.array[,,"b0"])
b1_mean <- mean(mcmc.array[,,"b1"])
b2_mean <- mean(mcmc.array[,,"b2"])
sigma_alpha_mean_2 <- mean(mcmc.array[,,"sigma_alpha"])

```

```{r,results='hide'}
"
model {
for( i in 1:n) {	
logit(p.i[i]) <- (
# different notation; using betas for main effects and b's for varying effects
  b0 + b1*c.dist100[i] + b2*c.log.arsenic[i] + alpha.j[getj1.i[i]]
  )
  y.i[i] ~ dbern(p.i[i])
}
for (j in 1:J){
 alpha.j[j] ~ dnorm(0, 1/sigma_alpha^2)
}
b0  ~ dnorm(0.0,1.0E-4)
b1 ~ dnorm(0.0,1.0E-4)
b2 ~ dnorm(0.0,1.0E-4)
sigma_alpha ~ dunif(0,200)

} # end model

"

```
We can see that the variance of the first group mean is smaller than that of the second group mean

```{r}
print ("Grouping 1 sigma alpha")
print (sigma_alpha_mean_1)
print ("Grouping 2 sigma alpha")
print (sigma_alpha_mean_2)

```

This make sense because we have reduced the variance within each group by putting households that are close to each other in the dataset , under the same village cluster. As long as there is any correlation between row i and row i+1 (such as survey collectors going from one town, to the town adjacent) this will reduce the variance within the cluster. 

#### b)

```{r}

invlogit <- function(x) {exp(x)/(1 + exp(x))}
p <- c()
p <- c(p,quantile(invlogit(mcmc.array[,,"b0"]+ (.5-mean(dist100))*mcmc.array[,,"b1"] +(log(1.2)-mean(log.arsenic))*mcmc.array[,,"b2"]  +mcmc.array[,,"alpha.j[1]"]),c(.025,.975)))

p <-  c(p, quantile(invlogit(mcmc.array[,,"b0"]+ (.5-mean(dist100))*mcmc.array[,,"b1"] +(log(1.2)-mean(log.arsenic))*mcmc.array[,,"b2"]  +mcmc.array[,,"alpha.j[2]"]),c(.025,.975)))

p <- c(p, 
quantile(invlogit(mcmc.array[,,"b0"]+ (.5-mean(dist100))*mcmc.array[,,"b1"] +(log(1.2)-mean(log.arsenic))*mcmc.array[,,"b2"]  +mcmc.array[,,"alpha.j[3]"]),c(.025,.975)))
p <- c(p,
quantile(invlogit(mcmc.array[,,"b0"]+ (.5-mean(dist100))*mcmc.array[,,"b1"] +(log(1.2)-mean(log.arsenic))*mcmc.array[,,"b2"]  +mcmc.array[,,"alpha.j[4]"]),c(.025,.975)))

p <- c(p,
quantile(invlogit(mcmc.array[,,"b0"]+ (.5-mean(dist100))*mcmc.array[,,"b1"] +(log(1.2)-mean(log.arsenic))*mcmc.array[,,"b2"]  +mcmc.array[,,"alpha.j[5]"]),c(.025,.975)))

p <- matrix(p,ncol=2,byrow = TRUE)

df_p <- data.frame(p)
rownames(df_p)=c("group1","group2","group3","group4","group5")
colnames(df_p)=c("L","U")

print (df_p)


```